---
title: "Martingale Posterior Distributions for Time Series Models"
---

This code generates figures for the article 
*Martingale Posterior Distributions for Time Series Models*.

Set image properties and load necessary libraries:

```{r setup}
knitr::opts_chunk$set(dev = 'png', fig.path = "images/", dpi = 300,
                      fig.asp = .66, echo = TRUE)

library(MARSS)
library(foreach)
library(doParallel)
library(parallel)
library(tidyverse)

set.seed(8080)
```

Figure 1: Martingale posterior for transition matrix of a discrete hidden Markov
model.

```{r fig1-code}
n <- 50
N <- 1000
M <- 1000
s <- c(1, 2)
x <- numeric(n)
x[1] <- sample(s, 1)
Q <- matrix(c(0.2, 0.6, 0.8, 0.4), nrow = 2)
Qhat <- matrix(c(0, 0, 0, 0), nrow = 2)
for (i in 2:n) {
  x[i] <- sample(s, 1, prob = Q[x[i - 1], ])
  Qhat[x[i - 1], x[i]] <- Qhat[x[i - 1], x[i]] + 1
}
Qpost <- array(NA, dim = c(M, 2, 2))
for (j in 1:M) {
  QQ <- Qhat
  xx <- numeric(N + 1)
  xx[1] <- x[n]
  for (i in 2:(N + 1)) {
    xx[i] <- sample(s, 1, prob = QQ[xx[i - 1], ])
    QQ[xx[i - 1], xx[i]] <- QQ[xx[i - 1], xx[i]] + 1
  }
  for (ss in s) {
    QQ[ss, ] <- QQ[ss, ] / sum(QQ[ss, ])
  }
  Qpost[j, , ] <- QQ
}
for (ss in s) {
  Qhat[ss, ] <- Qhat[ss, ] / sum(Qhat[ss, ])
}
Qhat
apply(Qpost, 2:3, mean)
```

```{r fig1-plot}
(tibble(`P(1 -> 1)` = Qpost[, 1, 1], `P(2 -> 2)` = Qpost[, 2, 2]) %>% 
  ggplot(aes(x = `P(1 -> 1)`, y = `P(2 -> 2)`)) +
  geom_point(color = 'grey50') +
  geom_vline(xintercept = Qhat[1, 1], linetype = 'dashed') +
  geom_hline(yintercept = Qhat[2, 2], linetype = 'dashed') +
  geom_point(x = Q[1, 1], y = Q[2, 2], shape = 4, size = 3, color = 'red') +
  theme_bw()) %>% 
  ggExtra::ggMarginal()
```

Figure 2: Comparison of Bayesian and Martingale posteriors for autoregressive 
time series.

```{r fig2-code}
n <- 20
N <- 500
M <- 1000
theta <- 0.6
sig <- 1

x <- numeric(n)
for (i in 2:n) {
  x[i] <- theta * x[i - 1] + rnorm(1, sd = sqrt(sig))
}
thhat <- sum(x[2:n] * x[1:(n - 1)]) / sum(x[1:(n - 1)] ^ 2)

thpost <- numeric(M)
for (j in 1:M) {
  th <- thhat
  sm <- sum(x[1:(n - 1)] ^ 2)
  xx <- c(x, numeric(N))
  for (i in 1:N) {
    xx[n + i] <- th * xx[n + i - 1] + rnorm(1, sd = sqrt(sig))
    th <- (xx[n + i] * xx[n + i - 1] + sm * th) / (sm + xx[n + i - 1] ^ 2)
    sm <- sm + xx[n + i - 1] ^ 2
  }
  thpost[j] <- th
}
```

```{r fig2-plot}
ggplot(data = tibble(Theta = thpost), aes(x = Theta)) + 
  geom_histogram(aes(y = ..density..), bins = 20, color = 'black', fill = NA) +
  geom_function(fun = dnorm,
                args = list(mean = thhat,
                            sd = sqrt(sig / sum(x[1:(n - 1)] ^ 2))),
                linetype = 'dashed',
                color = 'red') +
  ylab('Density') +
  theme_bw()
```

Figure 3: Martingale posterior for a Poisson autoregressive model

```{r fig3-code}
n <- 20
N <- 500
M <- 1000
theta <- 1.5
r <- 1
x <- numeric(n)
x[1] <- rgamma(1, theta, 1)
for (i in 2:n) {
  x[i] <- rgamma(1, theta + rpois(1, r * x[i - 1]), 1 + r)
}
thhat <- mean(x)
thpost <- numeric(M)
for (j in 1:M) {
  th <- thhat
  xx <- c(x, numeric(N))
  for (i in 1:N) {
    xx[n + i] <- rgamma(1, th + rpois(1, r * xx[n + i - 1]), 1 + r)
    th <- ((n + i - 1) * th + xx[n + i]) / (n + i)
  }
  thpost[j] <- th
}
```

```{r fig3-plot}
ggplot(data = tibble(Theta = thpost), aes(x = Theta)) + 
  geom_histogram(aes(y = ..density..), bins = 20, color = 'black', fill = NA) +
  geom_vline(xintercept = thhat, linetype = 'dashed') +
  geom_vline(xintercept = theta, linetype = 'dashed', color = 'red') +
  ylab('Density') +
  theme_bw()
```

Figure 4: Gradient ascent posterior for an ARCH model.

```{r fig4-code}

```

```{r fig4-plot}

```

Load in `graywhales` data, minor preprocessing, set forecast parameters, fit 
initial model.

```{r load-in}
data(graywhales)

clrs <- c(mle = '#882255', mrtn = '#44AA99', bays = '#332288',
          bays1 = '#88CCEE', bays2 = '#117733', boot = '#CC6677')

years_ <- graywhales[,1]
years <- seq(min(years_), max(years_))
loggraywhales <- rep(NA, length(years))
loggraywhales[years %in% years_] <- log(graywhales[,2])

n <- length(loggraywhales)
Hzn <- 25
N <- 500
M <- 100 * detectCores()

kem <- MARSS(loggraywhales)
```

Generate a forecast from the initial model.

Figure 5: Initial MARSS model fit.

```{r model0, cache = TRUE}
cl <- makeCluster(detectCores(), type = 'PSOCK')
clusterExport(cl, c('MARSS'))
registerDoParallel(cl)

fcst0 <- foreach (m = 1:M) %dopar% {
    synthx <- numeric(Hzn)
    synthy <- numeric(Hzn)
    mod <- kem
    xn <- predict(mod, n.ahead = 0, type = 'xtT')$pred[n, 'estimate']
    return(xn + cumsum(rnorm(Hzn, coef(mod)$U, sqrt(coef(mod)$Q))))
}

stopCluster(cl)
```

```{r model0-plot}
fcst_mod0 <- tibble(Year = rep(1951 + 1:n, M),
                 LogCount = rep(predict(kem, type = 'xtT')$pred$estimate[1:n],
                                M),
                 Data = rep(loggraywhales, M),
                 synth = rep(1:M, each = length(loggraywhales)))
fcst_mod1 <- tibble(Year = 1951 + n + rep(1:Hzn, M),
                    LogCount = unlist(fcst0),
                    Data = NA,
                    synth = rep(1:M, each = Hzn))
mod0_plot <- fcst_mod0 %>% 
    ggplot(aes(x = Year, y = LogCount, group = synth)) +
    geom_line(size = 0.5, alpha = 0.2) +
    geom_point(aes(y = Data), color = 'white', size = 2) +
    geom_point(aes(y = Data), color = 'black') +
    theme_bw() +
    theme(legend.position = 'none')
mod0_plot
```

Obtain Martingale posterior from MARSS model.

```{r, cache = TRUE}
comb1 <- function(out1, out2) {
  mods <- append(out1[[4]], out2[[4]])
  synthy <- cbind(out1[[3]], out2[[3]])
  synthx <- cbind(out1[[2]], out2[[2]])
  pred <- cbind(out1[[1]], out2[[1]])
  coef_seq <- cbind(out1[5], out2[5])

  return(list(pred = pred, synthx = synthx, synthy = synthy, mods = mods,
              coef_seq = coef_seq))
}


cl <- makeCluster(detectCores(), type = 'PSOCK')
clusterExport(cl, c('MARSS'))
registerDoParallel(cl)

t <- (res <- foreach (m = 1:M,
                      .combine = 'comb1') %dopar% {
  synthx <- numeric(N)
  synthy <- numeric(N)
  coef_seq <- list()
  xn <- predict(kem, n.ahead = 0, type = 'xtT')$pred[length(loggraywhales),
                                                     'estimate']
  for (n in 1:N) {
    if (n == 1) {
      mod <- kem
    } else {
      mod <- MARSS(c(loggraywhales, synthy[1:(n - 1)]),
                   inits = coef(mod),
                   silent = TRUE,
                   control = list(minit = 1,
                                  min.iter.conv.test = 1))
    }
    xn <- xn + coef(mod)$U + rnorm(1, 0, sqrt(coef(mod)$Q))
    synthx[n] <- xn
    synthy[n] <- synthx[n] + rnorm(1, 0, sqrt(coef(mod)$R))
    coef_seq[[n]] <- as.numeric(coef(mod)[c('U', 'x0', 'R', 'Q')])
  }
  prd <- predict(mod, n.ahead = 1, type = 'xtT')
  pred <- prd$pred$estimate
  return(list(pred, synthx, synthy, list(mod), coef_seq))
}) %>% system.time()

stopCluster(cl)

print(t)
```

Some figures showing the synthetic data...

```{r}
df_y0 <- tibble(Year = 1951 + 1:n,
                LogCount = loggraywhales,
                synth = 0)
df_y <- tibble(Year = 1951 + n + rep(1:N, M),
               LogCount = as.numeric(res$synthy),
               synth = rep(1:M, each = N))
```

```{r}
df_y0 %>%
  rbind(df_y) %>%
  ggplot(aes(x = Year, y = LogCount, group = synth)) +
  geom_point(data = df_y0) +
  geom_line(data = df_y, size = 0.5, alpha = 0.2) +
  theme_bw() +
  theme(legend.position = 'none')
```

```{r}
df_x0 <- tibble(Year = rep(1951 + 1:n, M),
                LogCount = rep(predict(kem, type = 'xtT')$pred$estimate, M),
                synth = rep(1:M, each = n))
df_xt <- tibble(Year = 1951 + n + rep(1:N, M),
                LogCount = as.numeric(res$synthx),
                synth = rep(1:M, each = N))
df_x <- tibble(Year = 1951 + rep(1:(n + N), M),
               LogCount = as.numeric(res$pred),
               synth = rep(1:M, each = n + N))
```

```{r x_synth}
df_x0 %>%
  rbind(df_xt) %>%
  ggplot(aes(x = Year, y = LogCount, group = synth)) +
  geom_point(data = df_y0, color = 'white', size = 2) +
  geom_point(data = df_y0, color = 'black') +
  geom_point(data = df_y, color = 'white', size = 2, alpha = 0.1) +
  geom_point(data = df_y, color = 'black', alpha = 0.1) +
  geom_line(size = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = 'none')
```

```{r x_smooth}
df_x %>%
  ggplot(aes(x = Year, y = LogCount, group = synth)) +
  geom_point(data = df_y0) +
  geom_point(data = df_y, alpha = 0.1) +
  geom_line(size = 0.5, alpha = 0.5) +
  theme_bw() +
  theme(legend.position = 'none')
```

Figure 6: Trace plots.

```{r}
# Coefficient convergence
arr <- array(unlist(res$coef_seq), dim = c(4, N, M))
df_theta <- tibble(Iteration = rep(1:N, M),
                   Chain = rep(1:M, each = N),
                   u = arr[1, , ] %>% as.numeric,
                   x0 = arr[2, , ] %>% as.numeric,
                   r = arr[3, , ] %>% as.numeric,
                   q = arr[4, , ] %>% as.numeric)
```

```{r trace_u}
(df_theta %>%
    ggplot(aes(x = Iteration, y = u, group = Chain)) +
    geom_line(alpha = 0.2) +
    geom_hline(yintercept = coef(kem)$U, size = 1, color = clrs[['mle']]) +
    theme_bw() +
    geom_point(data = subset(df_theta, Iteration == N), alpha = 0)) %>%
  ggExtra::ggMarginal(margins = 'y')
```

```{r trace_x0}
(df_theta %>%
    ggplot(aes(x = Iteration, y = x0, group = Chain)) +
    geom_line(alpha = 0.2) +
    geom_hline(yintercept = coef(kem)$x0, size = 1, color = clrs[['mle']]) +
    theme_bw() +
    geom_point(data = subset(df_theta, Iteration == N), alpha = 0)) %>%
  ggExtra::ggMarginal(margins = 'y')
```

```{r trace_r}
(df_theta %>%
    ggplot(aes(x = Iteration, y = r, group = Chain)) +
    geom_line(alpha = 0.2) +
    geom_hline(yintercept = coef(kem)$R, size = 1, color = clrs[['mle']]) +
    theme_bw() +
    geom_point(data = subset(df_theta, Iteration == N), alpha = 0)) %>%
  ggExtra::ggMarginal(margins = 'y')
```

```{r trace_q}
(df_theta %>%
    ggplot(aes(x = Iteration, y = q, group = Chain)) +
    geom_line(alpha = 0.2) +
    geom_hline(yintercept = coef(kem)$Q, size = 1, color = clrs[['mle']]) +
    theme_bw() +
    geom_point(data = subset(df_theta, Iteration == N), alpha = 0)) %>%
  ggExtra::ggMarginal(margins = 'y')
```

Pair plot for Martingale posterior.

```{r corrplot, fig.asp = 1}
df_theta %>%
  subset(Iteration == N) %>%
  select(c('u', 'x0', 'r', 'q')) %>%
  GGally::ggpairs() +
  theme_bw()
```

Obtain martingale forecast, optional figure.

```{r, cache = TRUE}
comb2 <- function(out1, out2) {
  synthy <- cbind(out1[[3]], out2[[3]])
  synthx <- cbind(out1[[2]], out2[[2]])
  pred <- cbind(out1[[1]], out2[[1]])

  return(list(pred = pred, synthx = synthx, synthy = synthy))
}

cl <- makeCluster(detectCores(), type = 'PSOCK')
clusterExport(cl, c('MARSS'))
registerDoParallel(cl)

fcst <- foreach (m = 1:M,
                 .combine = 'comb2') %dopar% {
    mod <- res$mods[[m]]
    xn <- predict(mod, n.ahead = 0, type = 'xtT')$pred[n, 'estimate']
    synthx <- xn + cumsum(rnorm(Hzn, coef(mod)$U, sqrt(coef(mod)$Q)))
    synthy <- rnorm(Hzn, synthx, sqrt(coef(mod)$R))
    prd <- predict(mod, n.ahead = 1, type = 'xtT')
    pred <- prd$pred$estimate
    return(list(pred, synthx, synthy))
  }

stopCluster(cl)
```

```{r}
fcst_x0 <- tibble(Year = rep(1951 + 1:n, M),
                 LogCount = as.numeric(sapply(1:M, function(m) 
                   predict(res$mods[[m]], type = 'xtT')$pred$estimate[1:n])),
                 synth = rep(1:M, each = n))
fcst_xt <- tibble(Year = 1951 + n + rep(1:Hzn, M),
                  LogCount = as.numeric(fcst$synthx),
                  synth = rep(1:M, each = Hzn))
```

```{r fcst}
fcst_plot <- fcst_x0 %>%
    rbind(fcst_xt) %>%
    ggplot(aes(x = Year, y = LogCount, group = synth)) +
    geom_line(size = 0.5, alpha = 0.2) +
    # geom_point(data = df_y0) +
    # geom_point(data = df_y, alpha = 0.1) +
    # geom_point(data = subset(fcst_xt, Year == max(Year)), alpha = 0) +
    geom_point(data = tibble(Year = 1951 + 1:n, LogCount = loggraywhales, 
                             synth = NA),
               color = 'white', size = 2) +
    geom_point(data = tibble(Year = 1951 + 1:n, LogCount = loggraywhales, 
                             synth = NA),
               color = 'black') +
    theme_bw() +
    theme(legend.position = 'none')
fcst_plot
```

Obtain parametric bootstrap results.

```{r, cache = TRUE}
cl <- makeCluster(detectCores(), type = 'PSOCK')
clusterExport(cl, c('MARSS'))
registerDoParallel(cl)

res_pboot <- foreach (m = 1:M,
                      .combine = 'comb1') %dopar% {
  synthx <- numeric(length(loggraywhales))
  synthy <- numeric(length(loggraywhales))
  coef_seq <- list()
  synthx <- coef(kem)$x0 + cumsum(rnorm(n,
                                        coef(kem)$U,
                                        sqrt(coef(kem)$Q)))
  synthy <- rnorm(n, synthx, sqrt(coef(kem)$R))
  synthy[is.na(loggraywhales)] <- NA
  mod <- MARSS(synthy,
               inits = coef(kem),
               silent = TRUE)
  coef_seq[[1]] <- as.numeric(coef(mod)[c('U', 'x0', 'R', 'Q')])
  prd <- predict(mod, n.ahead = 1, type = 'xtT')
  pred <- prd$pred$estimate
  return(list(pred, synthx, synthy, list(mod), coef_seq))
}

stopCluster(cl)
```
```{r pboot-plot}
arr_pboot <- array(unlist(res_pboot$coef_seq), dim = c(4, M))
df_thpboot <- tibble(Chain = 1:M,
                     u = arr_pboot[1, ] %>% as.numeric,
                     x0 = arr_pboot[2, ] %>% as.numeric,
                     r = arr_pboot[3, ] %>% as.numeric,
                     q = arr_pboot[4, ] %>% as.numeric)
df_thpboot %>%
    select(c('u', 'x0', 'r', 'q')) %>%
    GGally::ggpairs() +
    theme_bw()
```

Figure 7: Forecast Comparison.

```{r fcst_compare}
rbind(fcst_mod1 %>% 
        filter(Year == max(Year)) %>% 
        select(LogCount) %>% 
        mutate(Method = '\U03B8 fixed at MLE'),
      fcst_xt %>% 
        filter(Year == max(Year)) %>% 
        select(LogCount) %>% 
        mutate(Method = '\U03B8 recurisvely updated')) %>% 
  ggplot(aes(x = LogCount, fill = Method)) +
  xlab(paste('Forecasted LogCount in', 1997 + Hzn)) +
  geom_density(color = NA, alpha = 0.7) +
  geom_vline(xintercept = predict(kem, type = 'xtT',
                                  n.ahead = Hzn)$pred[length(loggraywhales) + Hzn, 'estimate'],
             color = clrs[['mle']]) +
  geom_vline(xintercept = log(16650),
             color = 'black') +
  scale_fill_manual(values = c(clrs[['mle']], clrs[['mrtn']])) +
  ylab('Density') +
  theme_bw()
```

Obtain Bayesian results.

```{r stan}
scode <- "
data {
  int<lower=0> TT; // length of ts
  int<lower=0> N; // num of ts; rows of y
  int<lower=0> n_pos; // number of non-NA values in y
  int<lower=0> col_indx_pos[n_pos]; // col index of non-NA vals
  int<lower=0> row_indx_pos[n_pos]; // row index of non-NA vals
  vector[n_pos] y;
}
parameters {
  vector[N] x0; // initial states
  real u;
  vector[N] pro_dev[TT]; // refed as pro_dev[TT,N]
  real<lower=0> sd_q;
  real<lower=0> sd_r[N]; // obs variances are different
}
transformed parameters {
  vector[N] x[TT]; // refed as x[TT,N]
  for(i in 1:N){
    x[1,i] = x0[i] + u + pro_dev[1,i];
    for(t in 2:TT) {
      x[t,i] = x[t-1,i] + u + pro_dev[t,i];
    }
  }
}
model {
  sd_q ~ cauchy(0,5);
  for(i in 1:N){
    x0[i] ~ normal(y[i], 10); // assume no missing y[1]
    sd_r[i] ~ cauchy(0,5);
    for(t in 1:TT){
    pro_dev[t,i] ~ normal(0, sd_q);
    }
  }
  u ~ normal(0,2);
  for(i in 1:n_pos){
    y[i] ~ normal(x[col_indx_pos[i], row_indx_pos[i]], sd_r[row_indx_pos[i]]);
  }
}
generated quantities {
  vector[n_pos] log_lik;
  for (n in 1:n_pos) log_lik[n] = normal_lpdf(y[n] | x[col_indx_pos[n], row_indx_pos[n]], sd_r[row_indx_pos[n]]);
}
"
Y <- t(matrix(loggraywhales))
ypos <- Y[!is.na(Y)]
n_pos <- length(ypos)  # number on non-NA ys
indx_pos <- which(!is.na(Y), arr.ind = TRUE)  # index on the non-NAs
col_indx_pos <- as.vector(indx_pos[, "col"])
row_indx_pos <- as.vector(indx_pos[, "row"])
mod <- rstan::stan(
  model_code = scode,
  data = list(
    y = ypos,
    TT = ncol(Y),
    N = nrow(Y),
    n_pos = n_pos,
    col_indx_pos = col_indx_pos,
    row_indx_pos = row_indx_pos
  ),
  pars = c("sd_q", "x", "sd_r", "u", "x0"),
  chains = 3,
  iter = 4000,
  thin = 1
)
```

```{r}
pars <- rstan::extract(mod)
df_thetastan <- tibble(Iteration = 1:6000,
                       u = pars$u %>% as.numeric,
                       x0 = pars$x0 %>% as.numeric,
                       r = pars$sd_r %>% as.numeric,
                       q = pars$sd_q %>% as.numeric)
df_theta$mode <- 'Martingale'
df_thetastan$mode <- 'Bayesian'
df_thetastan$r <- df_thetastan$r ^ 2
df_thetastan$q <- df_thetastan$q ^ 2
```

Figure 8: Pairplot for Bayes, Bootstrap, and Martingale results.

```{r corrplot2, fig.asp = 1}
p_names <- c('u', 'x0', 'r', 'q')
pairs <- rbind(df_thetastan[sample(2000:6000, M), ] %>%
        select(c(p_names, 'mode')),
      df_thpboot %>%
        mutate(mode = 'Bootstrap') %>%
        select(c(p_names, 'mode')),
      df_theta %>%
        subset(Iteration == N) %>%
        select(c(p_names, 'mode'))) %>%
  GGally::ggpairs(columns = 1:4, aes(color = mode),
                  diag = list(continuous = GGally::wrap('densityDiag',
                                                        alpha = 0.7,
                                                        color = NA)),
                  lower = list(continuous = GGally::wrap('points',
                                                         alpha = 0.2,
                                                         shape = 18)),
                  upper = list(continuous = GGally::wrap('cor'))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
th_mle <- as.numeric(coef(kem)[c('U', 'x0', 'R', 'Q')])
for(i in 1:pairs$nrow) {
  for(j in 1:pairs$ncol){
    if (i == j) {
      pairs[i,j] <- pairs[i,j] +
        geom_vline(xintercept = mean(df_thetastan[, p_names[i]][[1]]), 
                  color = clrs[['bays']], alpha = 0.5) +
        geom_vline(xintercept = mean(df_thpboot[, p_names[i]][[1]]), 
                  color = clrs[['boot']], alpha = 0.5) +
        geom_vline(xintercept = mean(df_theta[df_theta$Iteration == N,
                                              p_names[i]][[1]]), 
                  color = clrs[['mrtn']], alpha = 0.5) +
        geom_vline(xintercept = th_mle[i], color = clrs[['mle']])
    }
        pairs[i,j] <- pairs[i,j] + 
          scale_fill_manual(values = c(clrs[['bays']], clrs[['boot']],
                                       clrs[['mrtn']]))
        pairs[i,j] <- pairs[i,j] + 
          scale_color_manual(values = c(clrs[['bays']], clrs[['boot']],
                                        clrs[['mrtn']]))
  }
}
pairs
```

Obtain some other Bayesian results.

```{r}
scode_ <- "
data {
  int<lower=0> TT; // length of ts
  int<lower=0> N; // num of ts; rows of y
  int<lower=0> n_pos; // number of non-NA values in y
  int<lower=0> col_indx_pos[n_pos]; // col index of non-NA vals
  int<lower=0> row_indx_pos[n_pos]; // row index of non-NA vals
  vector[n_pos] y;
}
parameters {
  vector[N] x0; // initial states
  real u;
  vector[N] pro_dev[TT]; // refed as pro_dev[TT,N]
  real<lower=0> sd_q;
  real<lower=0> sd_r[N]; // obs variances are different
}
transformed parameters {
  vector[N] x[TT]; // refed as x[TT,N]
  for(i in 1:N){
    x[1,i] = x0[i] + u + pro_dev[1,i];
    for(t in 2:TT) {
      x[t,i] = x[t-1,i] + u + pro_dev[t,i];
    }
  }
}
model {
  sd_q ~ cauchy(0,5);
  for(i in 1:N){
"
scode1 <- "
    x0[i] ~ cauchy(y[i], 10); // assume no missing y[1]
"
scode2 <- "
    x0[i] ~ normal(y[i], sqrt(sd_q ^ 2 + sd_r[i] ^ 2)); // assume no missing y[1]
"
scode__ <- "
    sd_r[i] ~ cauchy(0,5);
    for(t in 1:TT){
    pro_dev[t,i] ~ normal(0, sd_q);
    }
  }
  u ~ normal(0,2);
  for(i in 1:n_pos){
    y[i] ~ normal(x[col_indx_pos[i], row_indx_pos[i]], sd_r[row_indx_pos[i]]);
  }
}
generated quantities {
  vector[n_pos] log_lik;
  for (n in 1:n_pos) log_lik[n] = normal_lpdf(y[n] | x[col_indx_pos[n], row_indx_pos[n]], sd_r[row_indx_pos[n]]);
}
"
Y <- t(matrix(loggraywhales))
ypos <- Y[!is.na(Y)]
n_pos <- length(ypos)  # number on non-NA ys
indx_pos <- which(!is.na(Y), arr.ind = TRUE)  # index on the non-NAs
col_indx_pos <- as.vector(indx_pos[, "col"])
row_indx_pos <- as.vector(indx_pos[, "row"])
mod1 <-
  rstan::stan(
    model_code = paste0(scode_, scode1, scode__),
    data = list(
      y = ypos,
      TT = ncol(Y),
      N = nrow(Y),
      n_pos = n_pos,
      col_indx_pos = col_indx_pos,
      row_indx_pos = row_indx_pos
    ),
    pars = c("sd_q", "x", "sd_r", "u", "x0"),
    chains = 3,
    iter = 4000,
    thin = 1
  )
mod2 <-
  rstan::stan(
    model_code = paste0(scode_, scode2, scode__),
    data = list(
      y = ypos,
      TT = ncol(Y),
      N = nrow(Y),
      n_pos = n_pos,
      col_indx_pos = col_indx_pos,
      row_indx_pos = row_indx_pos
    ),
    pars = c("sd_q", "x", "sd_r", "u", "x0"),
    chains = 3,
    iter = 4000,
    thin = 1
  )
```

```{r}
df_thetastan$mode <- 'Uninformed'
pars <- rstan::extract(mod1)
df_thetastan1 <- tibble(Iteration = 1:6000,
                        u = pars$u %>% as.numeric,
                        x0 = pars$x0 %>% as.numeric,
                        r = pars$sd_r %>% as.numeric,
                        q = pars$sd_q %>% as.numeric)
df_thetastan1$mode <- 'Uninformed+'
df_thetastan1$r <- df_thetastan1$r ^ 2
df_thetastan1$q <- df_thetastan1$q ^ 2
pars <- rstan::extract(mod2)
df_thetastan2 <- tibble(Iteration = 1:6000,
                        u = pars$u %>% as.numeric,
                        x0 = pars$x0 %>% as.numeric,
                        r = pars$sd_r %>% as.numeric,
                        q = pars$sd_q %>% as.numeric)
df_thetastan2$mode <- 'Hierarchical'
df_thetastan2$r <- df_thetastan2$r ^ 2
df_thetastan2$q <- df_thetastan2$q ^ 2
```

Figure 9: Pair plot for three Bayesian priors.

```{r, fig.asp = 1}
pairs <- rbind(df_thetastan1[sample(2000:6000, M), ] %>%
               select(c('u', 'x0', 'r', 'q', 'mode')),
               df_thetastan[sample(2000:6000, M), ] %>%
               select(c('u', 'x0', 'r', 'q', 'mode')),
               df_thetastan2[sample(2000:6000, M), ] %>%
               select(c('u', 'x0', 'r', 'q', 'mode'))
               ) %>%
  GGally::ggpairs(columns = 1:4, aes(color = mode),
                  diag = list(continuous = GGally::wrap('densityDiag',
                                                        alpha = 0.7,
                                                        color = NA)),
                  lower = list(continuous = GGally::wrap('points',
                                                         alpha = 0.2,
                                                         shape = 18)),
                  upper = list(continuous = GGally::wrap('cor'))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
th_mle <- as.numeric(coef(kem)[c('U', 'x0', 'R', 'Q')])
for(i in 1:pairs$nrow) {
  for(j in 1:pairs$ncol){
    if (i == j) {
      pairs[i,j] <- pairs[i,j] +
        geom_vline(xintercept = mean(df_thetastan[, p_names[i]][[1]]), 
                  color = clrs[['bays']], alpha = 0.5) +
        geom_vline(xintercept = mean(df_thetastan1[, p_names[i]][[1]]), 
                  color = clrs[['bays1']], alpha = 0.5) +
        geom_vline(xintercept = mean(df_thetastan2[, p_names[i]][[1]]), 
                  color = clrs[['bays2']], alpha = 0.5) +
        geom_vline(xintercept = th_mle[i], color = clrs[['mle']])
    }
        pairs[i,j] <- pairs[i,j] + 
          scale_fill_manual(values = c(clrs[['bays2']], clrs[['bays']],
                                       clrs[['bays1']]))
        pairs[i,j] <- pairs[i,j] + 
          scale_color_manual(values = c(clrs[['bays2']], clrs[['bays']],
                                        clrs[['bays1']]))
  }
}
pairs
```
